TGI-spinal-power

### 2022-05-03 analysis/tgispine_lme_power.Rmd
formalising power analysis to determine how many observations per variable combination is requried

Create a model with more data points per variable, to extract more power
# First, let's try N = 100
```{r, include = FALSE}
# add more participants to the model, for simulation
model.coldN <- extend(model.cold, within = 'manipulation+condition+cold_location', n=100)
summary(model.coldN)
# check that number of data points per variable does = 30
dat_model <- xtabs(~ manipulation + condition + cold_location, data=getData(model.coldN))
print(dat_model)
```

# Run power analysis!
````{r}
# extract fixed effects from the model - take the lowest effect size from the model
tgi_effect <- fixef(model.coldN)["cold_locationprox_caud"]
# need to estimate minimum desired effect from the change in slope - considering values can range from 0-100, it might be simpler to think of these effect sizes as a proportion of 100 instead
# tgi_effect = -17.05 would perhaps leave an expected effect size of .17
# change the effect size to effect size of interest
fixef(model.cold)["manipulationTGI"] <- tgi_effect

# then run power analysis
powerSim(model.cold)
```

Now to run a power curve on the data to find out the optimal number of participants needed for the effect of interest
```{r}
# let's see if this works because it hasn't before
pc.cold <- powerCurve(model.coldN)
print(pc.cold)
```

Now make a second, more complicated model where cold ratings are regressed by both manipulation and dermatome position (condition)
```{r}
model.cold2 <- lmer(VAScold ~ manipulation + condition + (1+manipulation|ID), 
                    data = df_phs)
summary(model.cold2)
```

Finally a third model where all predictors are added - manipulation, condition and cold location
Still just on cold ratings
```{r}
model.cold3 <- lmer(cold ~ manipulation + condition + cold_location +
                      (1|ID), 
                    data = df_phs)
summary(model.cold3)
```

>>>>

Create a model with more data points per variable, to extract more power
```{r}
# First, let's try to add 30 observations per variable combination, which would mean N = 30
model.coldN <- extend(model.cold, within = 'manipulation+condition+cold_location', n=30)
summary(model.coldN)
# check that number of data points per variable does = 30
dat_model <- xtabs(~ manipulation + condition + cold_location, data=getData(model.coldN))
print(dat_model)
```

# Run power analysis!
````{r, results = 'hide'}
# extract fixed effects from the model - take the lowest effect size from the model
tgi_effect <- fixef(model.coldN)["manipulationTGI"]
cond_effect <- fixef(model.coldN)["conditionwithin"]
cold_loc_effect <- fixef(model.coldN)["cold_locationprox_caud"]
# need to estimate minimum desired effect from the change in slope - considering values can range from 0-100, it might be simpler to think of these effect sizes as a proportion of 100 instead
# tgi_effect = -17.05 would perhaps leave an expected effect size of .17
# change the effect size to effect size of interest
fixef(model.coldN)["cold_locationprox_caud"] <- cold_loc_effect

# then run power analysis
pwr.cold <- powerSim(model.coldN)
print(pwr.cold)
```

Now to run a power curve on the data to find out the optimal number of participants needed for the effect of interest
```{r}
# can run a power curve to determine the point of trade off between N observations and power
pc.cold <- powerCurve(model.coldN, within='manipulation+condition+cold_location', 
                      breaks= seq(0, 30, by = 3))
print(pc.cold)
```

# Plot power curve!
```{r}

```
